---
title: "Project 1 NBA"
output: html_notebook
---
```{r}
"Import reticulate to work between R and python as well as other packages"

library(reticulate)
use_virtualenv("r-reticulate")
py_available(TRUE)


```


```{python}
import math
import pandas as pd
import matplotlib.pyplot as plt
import numpy as np
from matplotlib import colors
from matplotlib.ticker import PercentFormatter

team_Data = pd.read_csv("../Resources/teamData.csv")
player_Data = pd.read_csv("../Resources/playerData.csv")
crossReference_Data = pd.read_csv("../Resources/crossReference.csv")

merged_player_data = player_Data.loc[(player_Data['Year'] >= 1986) & (player_Data['Year'] < 2018)]

#merged_player_data = merged_player_data[merged_player_data['Year']==year]
```


```{python}
selected_columns = ['Player', 'PER', 'TS%', 'ORB%', 'DRB%', 'AST%', 'STL%', 'BLK%', 'TOV%', 'WS', '3P%', '2P%', 'FT%', 'PTS', 'MP']
selected_columns_with_age = ['Player', 'PER', 'TS%', 'ORB%', 'DRB%', 'AST%', 'STL%', 'BLK%', 'TOV%', 'WS', '3P%', '2P%', 'FT%', 'PTS', 'MP', 'Age']

subset2py = merged_player_data[selected_columns]
subset_with_age = merged_player_data[selected_columns_with_age]
  subset2py2 = subset2py

Q1=subset2py2.quantile(0.25)
Q3=subset2py2.quantile(0.75)
IQR=Q3-Q1

 #set up quantiles and IQR with age
Q1a = subset_with_age.quantile(0.25)
Q3a = subset_with_age.quantile(0.75)
IQRa = Q3a-Q1a

subset2py_no_outlier = subset2py2[~((subset2py2<(Q1-1.5*IQR)))]
```

```{python}
percentile_df = pd.DataFrame()
for column in selected_columns[1:]:
  percentile_df[f"{column}_percentile"] = merged_player_data[column].rank(pct = True)
  percentile_df['Player Name'] = merged_player_data['Player']
  
percentile_df = percentile_df.dropna().reset_index()

main_perc_list = []


for index in range(0,percentile_df.shape[0]):
  row_sum = sum(list(percentile_df.iloc[index])[1:-2])
  row_main_perc = (row_sum/15)
  main_perc_list.append(row_main_perc)

main_percentile = pd.DataFrame(main_perc_list)
percentile_df['Main Percentile'] = main_percentile
[row, cols] = percentile_df.shape

```



```{python}

```


```{python}

```

```{python}

```

```{r}
"a list of varibles to intergrate into the regression"

"dependent varible: minutes played"
"independent varibles: 
  PER (player efficiency) 
  TS (true shooting %) 
  ORB (offensive rebound rate) 
  DRB (defensive rebound rate)
  AST (assist %)
  STL (steal %)
  BLK (Block %)
  TOV (turnover %)
  WS (win shares)
  3P% (3-Point %)
  2P% (2-point field goal %)
  FT% (Free throw %)
  PTS (points)"


```


```{python}


```

```{python}
subset2py2['ln_mp'] = np.log(subset2py2['MP'])
subset2py2['ln_pts'] = np.log(subset2py2['PTS'])
```
```{r}
hist(py$subset2py2$MP)
hist(py$subset2py2$ln_mp)
hist(py$subset2py_no_outlier$ln_mp)
hist(py$subset2py2$PTS)
hist(py$subset2py2$ln_pts)
hist(py$subset2py_no_outlier$ln_pts)
```



```{python}
#set up quantiles and IQR
Q1=subset2py2.quantile(0.25)
Q3=subset2py2.quantile(0.75)
IQR=Q3-Q1

#set up quantiles and IQR with age
Q1a = subset_with_age.quantile(0.25)
Q3a = subset_with_age.quantile(0.75)
IQRa = Q3a-Q1a

subset_no_outlier_with_age = subset_with_age[~(subset_with_age<(Q1-1.5*IQR))]



"To get rid of outliers in subset df"
subset2py_no_outlier = subset2py2[~((subset2py2<(Q1-1.5*IQR)) | (subset2py2 > (Q3+1.5*IQR))).any(axis=1)]
subset2py_no_outlier

#plt.hist(subset2py_no_outlier['TS%'], bins= 'auto', color='#0504aa', alpha=0.7, rwidth= 0.85)

#plt.grid(axis='y', alpha=0.75)
#plt.xlabel('True Shot %')
#plt.ylabel('Frequency')
#plt.title('My Very Own Histogram')
#plt.text(23, 45, r'$\mu=15, b=3$')


```
```{python}
#modify values in subset with no outliers to reflect Percentile
percentile_df = pd.DataFrame()
for column in selected_columns[1:]:
  #fileter by year
  percentile_df[f"{column}_percentile"] = merged_player_data[column].rank(pct = True)
  
percentile_df['Player Name'] = merged_player_data['Player']
percentile_df = percentile_df.dropna().reset_index()




```


```{python}
#To make a new variable call main percentage which averages all values in each row

main_perc_list = []


for index in range(0,percentile_df.shape[0]):
  row_sum = sum(list(percentile_df.iloc[index])[1:-2])
  row_main_perc = (row_sum/15)
  main_perc_list.append(row_main_perc)
  
row_main_perc
main_percentile = pd.DataFrame(main_perc_list)
percentile_df['Main Percentile'] = main_percentile
[row, cols] = percentile_df.shape



percentile_df['Adjusted Percentile'].value_counts





```


```{python}
#Define multiple formulas in order to get the SD and adjusted percentile

#mean : list -> float
def mean(data):
  total = sum(data)
  return total / len(data)


#numerator : list -> float
def numerator(data):
  SD_mean = mean(data)
  total = 0 
  for val in data:
    total = total + ((val - SD_mean)**2)
  return total


#std : list -> float
def std(data):
  num =  numerator(data)
  den = len(data)
  return math.sqrt(num/den)


#new_formula : list -> list
def new_formula(data): 
  percentile = []
  std_fun = std(data)
  mean_fun = mean(data)
  for value in data:
    adjusted_value = (value - mean_fun)/std_fun
    percentile.append(adjusted_value)
  return percentile
  
adjusted_percentile_df  = pd.DataFrame(new_formula(percentile_df['Main Percentile']))

print(std(percentile_df['Main Percentile']))
print(mean(percentile_df['Main Percentile']))

```
```{python}
plt.hist(percentile_df['Adjusted Percentile'])
plt.show()
```


```{python}
percentile_df['Adjusted Percentile'] = adjusted_percentile_df

list_main_percential_name= ['Main Percentile', 'Player Name']
list_main_percential_name_df = percentile_df[list_main_percential_name]

```

```{python}
for values in list_main_percential_name_df['Player Name']:
  if list_main_percential_name_df['Player Name'].value_counts() > 10
  print(list_main_percential_name_df['Main Percentile'])


#list_main_percential_name_df['Player Name'].value_counts() 

player_name = list_main_percential_name_df.groupby('Player Name')
mean_main_percentile = player_name['Main Percentile'].mean()
mean_main_percentile_df = pd.DataFrame({'Player Name': mean_main_percentile.index, 'Main Percentile': mean_main_percentile.values})
print(mean_main_percentile_df)

```

```{python}

```

```{python}


```


```{r}


```

```{python}

  
```

```{python}
selected_columns = ['Player', 'PER', 'TS%', 'ORB%', 'DRB%', 'AST%', 'STL%', 'BLK%', 'TOV%', 'WS', '3P%', '2P%', 'FT%', 'PTS', 'MP']
selected_columns_with_age = ['Player', 'PER', 'TS%', 'ORB%', 'DRB%', 'AST%', 'STL%', 'BLK%', 'TOV%', 'WS', '3P%', '2P%', 'FT%', 'PTS', 'MP', 'Age']

subset2py = merged_player_data[selected_columns]

```

```{python}
hp1_subset_list = ['MP', '3P', '2P', 'FT', 'ORB', 'BLK', 'DRB', 'STL']


hp1_subset = merged_player_data[hp1_subset_list]

```

```{python}
Q1h=hp1_subset.quantile(0.25)
Q3h=hp1_subset.quantile(0.75)
IQRh=Q3h-Q1h

hp1_subset['ln_mp'] = np.log(hp1_subset['MP'])
hp1_subset['ln_ft'] = np.log(hp1_subset['FT'])
hp1_subset['ln_orb'] = np.log(hp1_subset['ORB'])
hp1_subset['ln_2p'] = np.log(hp1_subset['2P'])
hp1_subset['ln_3p'] = np.log(hp1_subset['3P'])
hp1_subset['ln_blk'] = np.log(hp1_subset['BLK'])
hp1_subset['ln_drb'] = np.log(hp1_subset['DRB'])
hp1_subset['ln_stl'] = np.log(hp1_subset['STL'])

hp1_subset_no = hp1_subset[~((hp1_subset<(Q1h-1.5*IQRh))) | ((hp1_subset>(Q3h+1.5*IQRh)))]

#hp1_subset_no = hp1_subset_no_na = hp1_subset_no.dropna().reset_index()

hp1_working_df = hp1_subset_no[hp1_subset_no.replace([np.inf, -np.inf], np.nan).notnull().all(axis=1)]
#hp1_subset_no_na_tf = hp1_subset_no_na.replace([np.inf, -np.inf], np.nan, inplace=True)
#hp1_subset_no_na_tf.dropna().reset_index()
```

```{python}
hp1_working_df.loc[(hp1_working_df!=0).any(axis=1)]
```

```{r}
hp1_working_df  <- py$hp1_working_df

reg_hp1 <- lm(hp1_working_df$ln_mp ~ hp1_working_df$ln_blk + hp1_working_df$ln_drb + hp1_working_df$ln_stl + hp1_working_df$ln_ft + hp1_working_df$ln_orb + hp1_working_df$ln_3p + hp1_working_df$ln_2p)
```


```{r}
summary(reg_hp1)
summary(reg_no_outlier)
options(max.print = 1000000)
options(scipen=999)
```

```{r}
hist(py$hp1_subset_no$ln_orb)
hp1_subset_no <- py$hp1_subset_no

lm(hp1_subset_no$ln_mp ~ hp1_subset_no$ln_2p + hp1_subset_no$ln_3p + hp1_subset_no$ln_orb + hp1_subset_list$ln_ft)
```

```{r}
regresion_no_outlier_df <- py$subset2py_no_outlier

hist(subset_no_outlier$PER)
hist(regresion_no_outlier_df$`TS%`)
hist(regresion_no_outlier_df$ln_mp)
```
```{r}
reg_no_outlier <- lm(regresion_no_outlier_df$ln_mp ~ regresion_no_outlier_df$ln_pts + regresion_no_outlier_df$PER + regresion_no_outlier_df$`TS%` + regresion_no_outlier_df$`ORB%` + regresion_no_outlier_df$`DRB%` + regresion_no_outlier_df$`AST%` + regresion_no_outlier_df$`STL%` + regresion_no_outlier_df$`BLK%` + regresion_no_outlier_df$`TOV%` + regresion_no_outlier_df$WS + regresion_no_outlier_df$`3P%` + regresion_no_outlier_df$`2P%` + regresion_no_outlier_df$`FT%`) 
summary(reg_no_outlier)



```
```{r}
lm(regresion_no_outlier_df$ln_mp ~ regresion_no_outlier_df$`TS%`)
reg_no_outlier
extract_coef <- summary(reg_no_outlier)$coefficients
extract_coef[ , 4]
hist(regresion_no_outlier_df$`TS%`)
boxplot(regresion_no_outlier_df$`TS%`)
hist(regresion_no_outlier_df$ln_mp)
extract_coef_exp <- exp(extract_coef[ , 1])
adjusted_reg_no_outlier <- ((extract_coef_exp)-1)*100
adjusted_reg_no_outlier
```




```{r}
summary(reg_no_outlier)
options(max.print = 1000000)
options(scipen=999)

```


```{r}
#we first exponentiate the coefficient of census to obtain exp(0.00055773)=1.000558. To calculate the percent change, we can subtract one from this number and multiply by 100. Thus, for a one unit increase in the average daily number of patients (census), the average length of stay (length) increases by 0.06 percent.
(exp(-0.))-1
-.1381314*100
#For a one unit increase in minutes played results in a 13% decrease in a players efficiency rating 
```
```{r}

```

```{r}
ln_minutes_played <- log(py$minutes_played)
hist(py$subset2py_no_outlier$`TS%`)
hist(py$merged_player_data$`TS%`)
hist(py$subset2py$PER)
count(py$subset2py_no_outlier$Player)
```

```{r}
subset_no_outlier_with_age <- py$subset_no_outlier_with_age

plot(subset_no_outlier_with_age$Age, subset_no_outlier_with_age$PER, main="Scatterplot Example",
   xlab="Age", ylab="PER ", pch=19)


```

```{r}
"regression"
main_reg <- lm(ln_minutes_played ~ py$player_eff + py$true_shot_perc + py$off_reb_rate + py$def_reb_rate + py$assist_perc + py$steal_perc + py$block_perc + py$turnover_perc + py$win_shares + py$three_point_perc + py$two_point_perc + py$free_throw_perc + py$points)
options(max.print = 1000000)
options(scipen=999)
summary(main_reg)

main_reg_no_outlier <- lm()


```

```{r}
merged_player_data <- py$merged_player_data
merged_player_data$ln_MP <- log(merged_player_data$MP)

out2=with(merged_player_data, 
          by(merged_player_data, merged_player_data$Year, function(merged_player_data) lm(merged_player_data$ln_MP ~ merged_player_data$PER + merged_player_data$`TS%` + merged_player_data$`ORB%` + merged_player_data$`DRB%` + merged_player_data$`AST%` + merged_player_data$`STL%` + merged_player_data$`BLK%` + merged_player_data$`TOV%` + merged_player_data$WS + merged_player_data$`3P%` + merged_player_data$`2P%` + merged_player_data$`FT%` + merged_player_data$PTS, data=merged_player_data)))


out2[1]
```

```{python}
raw_data_groupby_year = subset_no_outlier_with_age.groupby['Year']

```

```{r}
subset_with_age_no_outlier_df <- py$subset_no_outlier_with_age

sd(subset_with_age_no_outlier_df$PTS)
```



```{r}
out3=with(subset_with_age_no_outlier_df,
          by(subset_with_age_no_outlier_df, subset_with_age_no_outlier_df$Year, function(subset_with_age_no_outlier_df) data=subset_with_age_no_outlier_df))

out3[2]

```

```{python}
r.
```


```{r}
all(is.na(subset_no_outlier))
```


```{r}
outsum <- lapply(out2, summary)
outsum[10]
```


```{r}
subsetr <- merged_player_data

```




```{r}

```


```{r}
lm(py$player_eff ~ py$minutes_played)
plot(py$minutes_played, py$player_eff)
```


```{r}

"a list of varibles to intergrate into the regression"
"dependent varible: minutes played"
"independent varibles: 
  PER (player efficiency) 
  TS (true shooting %) 
  ORB (offensive rebound rate) 
  DRB (defensive rebound rate)
  AST (assist %)
  STL (steal %)
  BLK (Block %)
  TOV (turnover %)
  WS (win shares)
  3P% (3-Point %)
  2P% (2-point field goal %)
  FT% (Free throw %)
  PTS (points)"


```

```{r}
for year in range(1986, 2019):
  merged_player_data = player_Data.loc[(player_Data['Year'] >= 1986) & (player_Data['Year'] < 2018)]

  merged_player_data = merged_player_data[merged_player_data['Year']==year]

  selected_columns = ['Player', 'PER', 'TS%', 'ORB%', 'DRB%', 'AST%', 'STL%', 'BLK%', 'TOV%', 'WS', '3P%', '2P%', 'FT%', 'PTS', 'MP']
  selected_columns_with_age = ['Player', 'PER', 'TS%', 'ORB%', 'DRB%', 'AST%', 'STL%', 'BLK%', 'TOV%', 'WS', '3P%', '2P%', 'FT%', 'PTS', 'MP', 'Age']

  subset2py = merged_player_data[selected_columns]
  subset_with_age = merged_player_data[selected_columns_with_age]
  subset2py2 = subset2py

  Q1=subset2py2.quantile(0.25)
  Q3=subset2py2.quantile(0.75)
  IQR=Q3-Q1

    #set up quantiles and IQR with age
  Q1a = subset_with_age.quantile(0.25)
  Q3a = subset_with_age.quantile(0.75)
  IQRa = Q3a-Q1a

  subset2py_no_outlier = subset2py2[~((subset2py2<(Q1-1.5*IQR)))]

  percentile_df = pd.DataFrame()
  for column in selected_columns[1:]:
    percentile_df[f"{column}_percentile"] = merged_player_data[column].rank(pct = True)
    percentile_df['Player Name'] = merged_player_data['Player']
  
  percentile_df = percentile_df.dropna().reset_index()

  main_perc_list = []


  for index in range(0,percentile_df.shape[0]):
    row_sum = sum(list(percentile_df.iloc[index])[1:-2])
    row_main_perc = (row_sum/15)
    main_perc_list.append(row_main_perc)
    
  main_percentile = pd.DataFrame(main_perc_list)
  percentile_df['Main Percentile'] = main_percentile
  [row, cols] = percentile_df.shape

  def mean(data):
    total = sum(data)
    return total / len(data)


    #numerator : list -> float
  def numerator(data):
    SD_mean = mean(data)
    total = 0 
    for val in data:
      total = total + ((val - SD_mean)**2)
    return total


    #std : list -> float
  def std(data):
    num =  numerator(data)
    den = len(data)
    return math.sqrt(num/den)


    #new_formula : list -> list
  def new_formula(data): 
    percentile = []
    std_fun = std(data)
    mean_fun = mean(data)
    for value in data:
      adjusted_value = (value - mean_fun)/std_fun
      percentile.append(adjusted_value)
    return percentile
    
  adjusted_percentile_df  = pd.DataFrame(new_formula(percentile_df['Main Percentile']))

  

  percentile_df['Adjusted Percentile'] = adjusted_percentile_df

    
  percentile_df.to_csv(f'../Resources/yearly_csvs/{year}_data.csv')
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
